GLOBALIZER: A NOVEL SUPERCOMPUTER SOFTWARE SYSTEM FOR SOLVING TIME-CONSUMING GLOBAL OPTIMIZATION PROBLEMS

### 2016-08-07 2016/GOM/paper.tex
Add references section of GOM paper (WIP)

number of key monographs (Törn, A., Žilinskas, A., 1989), (Horst, R., Tuy, H., 1990), (Zhigljavsky, A. A., 1991), (Pintér, J. D., 1996), (Strongin, R. G., Sergeyev, Y. D., 2000), (Locatelli, M., Schoen, F., 2013), (Floudas, C. A., Pardalos, M. P., 2016), etc.

@@

attention – see, for example, (Censor, Y., Zenios, S. A., 1998), (Strongin, R. G., Sergeyev, Y. D., 2000), (Ciegis, R., Henty, D., Kågström, B., Zilinskas, J., 2009), (Luque, G., Alba, E., 2011), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).

@@

optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A., 2013).

@@

unambiguously used for dimensional reduction (see, for example, Strongin, R. G., Sergeyev, Y. D., 2000).

@@

corresponding minimized function to satisfy the uniform Hölder condition (see Strongin, R. G., 1978, Strongin, R. G., Sergeyev, Y. D., 2000) i.e.

@@

The algorithms for the numerical construction of the Peano curve approximations are presented in (Strongin, R. G., Sergeyev, Y. D., 2000).

@@

\section{The Parallelization Approach and Global Optimization Algorithms}
\label{sec:parallel}
Let us consider the parallelization methods used widely in the theoretical and
practical application of parallel computing within the context of global optimization problems:
\begin{itemize}
  \item The distribution of the search domain \(D\) among the available computing units
  (data parallelization scheme). In the case of optimization problems, this approach is
  insufficient, since in organizing such computations the subdomain, which contain the
  sought global minimum, will be processed by only one processor, and, therefore,
  all of the remaining processors would perform the excess computations.
  \item The parallelization of the computational schemes for optimization algorithms
  (task parallelization scheme). This approach is also insufficient since the direct
  computational costs for executing optimization algorithms are relatively low
  (the majority of computations in the course of a global search are represented by
  calculations of the optimized function values due to the initial assumption of
  considerable computational costs for such calculations).
  \item The parallelization of the computations executed in order to obtain values
  for the optimized function. This approach can be applied since the most computationally
  costly part of the global search process will be parallelized. However, this method is not
  featured by the generality (the development of parallelization methods is to be performed
  every time from scratch while solving each particular optimization problem).
\end{itemize}
\par
Within the framework of information-statistical theory, a general approach to
parallelization computations when solving global optimization problems has been
proposed --- \textit{the parallelism of computations is provided by means of simultaneously
computing the values of the minimized function \(\varphi(y)\) at several different
points within the search domain \(D\)} – see, for example, (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
This approach provides parallelizaion for the most costly part of computations in the global search process.
\par
The global optimization algorithms implemented in Globalizer will be described step-by-step below.
In Subsection \ref{subsec:corepar}, the core sequential multidimensional algorithm of global search (MAGS) will be presented.
In Subsection \ref{subsec:sharedpar}, a parallel generalization of the MAGS algorithm for parallel computations on
computer systems with shared memory will be described. In Subsection \ref{subsec:distribpar}, the scheme
for parallel computations by multiprocessor systems with distributed memory will be provided.

@@

The information-statistical theory of global search formulated in (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000)

@@

presented as follows --- see (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000).

@@

The conditions for the algorithm’s convergence described above were examined, for example, in (Strongin, R. G., Sergeyev, Y. D., 2000).

\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}
Modern supercomputer computational systems consist of plenty of computer nodes,
which include several multicore processors. At that, random access memory is
shared for the computing nodes --- the content of any memory element can be read (written)
by any computer cores available at any arbitrary moment. In most cases, shared memory is
uniform --- the time characteristics for accessing memory are the same for all computer
cores and for all memory elements.
\par
The following speculations can form the basis for selecting parallel computation
organization methods. As was mentioned above when describing the MAGS algorithm,
the characteristics \(R(i),1\leqslant i\leqslant k+1\), calculated according to (\ref{step3})
can be interpreted as some measures of the importance of intervals with respect to the location of
the global minimum point. Following this understanding, each successive global search
iteration is executed within the most important interval with the maximum value of the
characteristic. So far, in this case it is obvious how to select the other intervals for
organizing simultaneous computations of the minimized function values at several different
points within the search domain as well --- these should be the intervals with the next
magnitudes of characteristics.
\par
The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
computer systems with shared memory (PMAGS) was developed based the approach considered
above and is practically identical to the MAGS scheme --- the differences consist just in the
following set of rules --- see (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
\par
\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
\begin{equation}
\label{step4par}
R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
\end{equation}
and to select \(p\) intervals with the indices \(t_j,1\leqslant j\leqslant p\), with
the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
\par
\textit{Rule 5'}. To execute new trials (computations of values for the minimized
function \(f(x)\)) at the points \(x_{k+j},1\leqslant j\leqslant p\), located in the
intervals with the highest characteristics from (\ref{step4par})
\begin{equation}
\label{step5par}
\begin{array}{cr}\\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\quad 1<t_j<k+1 \\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1
\end{array}
\end{equation}
\par
The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
intervals from (\ref{step4par}), in which the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leqslant j\leqslant p
\end{equation}
As before, if the stop condition is not fulfilled, the index \(k\) is
incremented by \(p\), and a new global search iteration is executed.
\par
The conditions for the developed parallel algorithm’s convergence and for the
non-redundancy of parallel computations have been examined in (Strongin, R. G., Sergeyev, Y. D., 2000).
Thus, in particular, when the conditions for convergence are satisfied, the parallel
computations are non-redundant as compared to the sequential method when using up to
\(2N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).

@@

instead of applying a single Peano curve \(y(x)\) has been proposed in (Strongin, R. G., 1992, Strongin, R. G., Sergeyev, Y. D., 2000).

@@

>>>>

number of key monographs \cite{zilinskTornGO}, \cite{horstTuyGO}, \cite{zhigljavskyRandGO},
\cite{pinterGO}, \cite{strSergGO}, \cite{locatelliSchoenGO}, \cite{floudasPardGO}, etc.

@@

attention – see, for example, (Censor, Y., Zenios, S. A., 1998), \cite{strSergGO}, (Ciegis, R., Henty, D., Kågström, B., Zilinskas, J., 2009), (Luque, G., Alba, E., 2011), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).

@@

optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R. G., 1978), \cite{strSergGO}, (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A., 2013).

@@

unambiguously used for dimensional reduction (see, for example, \cite{strSergGO}).

@@

corresponding minimized function to satisfy the uniform Hölder condition (see Strongin, R. G., 1978, \cite{strSergGO}) i.e.

@@

The algorithms for the numerical construction of the Peano curve approximations are presented in \cite{strSergGO}.

@@

\section{The Parallelization Approach and Global Optimization Algorithms}
\label{sec:parallel}
Let us consider the parallelization methods used widely in the theoretical and
practical application of parallel computing within the context of global optimization problems:
\begin{itemize}
  \item The distribution of the search domain \(D\) among the available computing units
  (data parallelization scheme). In the case of optimization problems, this approach is
  insufficient, since in organizing such computations the subdomain, which contain the
  sought global minimum, will be processed by only one processor, and, therefore,
  all of the remaining processors would perform the excess computations.
  \item The parallelization of the computational schemes for optimization algorithms
  (task parallelization scheme). This approach is also insufficient since the direct
  computational costs for executing optimization algorithms are relatively low
  (the majority of computations in the course of a global search are represented by
  calculations of the optimized function values due to the initial assumption of
  considerable computational costs for such calculations).
  \item The parallelization of the computations executed in order to obtain values
  for the optimized function. This approach can be applied since the most computationally
  costly part of the global search process will be parallelized. However, this method is not
  featured by the generality (the development of parallelization methods is to be performed
  every time from scratch while solving each particular optimization problem).
\end{itemize}
\par
Within the framework of information-statistical theory, a general approach to
parallelization computations when solving global optimization problems has been
proposed --- \textit{the parallelism of computations is provided by means of simultaneously
computing the values of the minimized function \(\varphi(y)\) at several different
points within the search domain \(D\)} – see, for example, \cite{strSergGO}, (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
This approach provides parallelizaion for the most costly part of computations in the global search process.
\par
The global optimization algorithms implemented in Globalizer will be described step-by-step below.
In Subsection \ref{subsec:corepar}, the core sequential multidimensional algorithm of global search (MAGS) will be presented.
In Subsection \ref{subsec:sharedpar}, a parallel generalization of the MAGS algorithm for parallel computations on
computer systems with shared memory will be described. In Subsection \ref{subsec:distribpar}, the scheme
for parallel computations by multiprocessor systems with distributed memory will be provided.

@@

The information-statistical theory of global search formulated in (Strongin, R. G., 1978), \cite{strSergGO}

@@

presented as follows --- see (Strongin, R. G., 1978), \cite{strSergGO}.

@@

The conditions for the algorithm’s convergence described above were examined, for example, in \cite{strSergGO}.

\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}
Modern supercomputer computational systems consist of plenty of computer nodes,
which include several multicore processors. At that, random access memory is
shared for the computing nodes --- the content of any memory element can be read (written)
by any computer cores available at any arbitrary moment. In most cases, shared memory is
uniform --- the time characteristics for accessing memory are the same for all computer
cores and for all memory elements.
\par
The following speculations can form the basis for selecting parallel computation
organization methods. As was mentioned above when describing the MAGS algorithm,
the characteristics \(R(i),1\leqslant i\leqslant k+1\), calculated according to (\ref{step3})
can be interpreted as some measures of the importance of intervals with respect to the location of
the global minimum point. Following this understanding, each successive global search
iteration is executed within the most important interval with the maximum value of the
characteristic. So far, in this case it is obvious how to select the other intervals for
organizing simultaneous computations of the minimized function values at several different
points within the search domain as well --- these should be the intervals with the next
magnitudes of characteristics.
\par
The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
computer systems with shared memory (PMAGS) was developed based the approach considered
above and is practically identical to the MAGS scheme --- the differences consist just in the
following set of rules --- see \cite{strSergGO}, (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
\par
\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
\begin{equation}
\label{step4par}
R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
\end{equation}
and to select \(p\) intervals with the indices \(t_j,1\leqslant j\leqslant p\), with
the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
\par
\textit{Rule 5'}. To execute new trials (computations of values for the minimized
function \(f(x)\)) at the points \(x_{k+j},1\leqslant j\leqslant p\), located in the
intervals with the highest characteristics from (\ref{step4par})
\begin{equation}
\label{step5par}
\begin{array}{cr}\\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\quad 1<t_j<k+1 \\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1
\end{array}
\end{equation}
\par
The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
intervals from (\ref{step4par}), in which the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leqslant j\leqslant p
\end{equation}
As before, if the stop condition is not fulfilled, the index \(k\) is
incremented by \(p\), and a new global search iteration is executed.
\par
The conditions for the developed parallel algorithm’s convergence and for the
non-redundancy of parallel computations have been examined in \cite{strSergGO}.
Thus, in particular, when the conditions for convergence are satisfied, the parallel
computations are non-redundant as compared to the sequential method when using up to
\(2N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).

@@

instead of applying a single Peano curve \(y(x)\) has been proposed in (Strongin, R. G., 1992, \cite{strSergGO}).

@@

\bibliographystyle{gOMS}
\bibliography{paper}
